Adaptive Disk Storage for Interaction Graphs

### 2014-04-09 notes/graph-adapt/main.tex
initial formalization for cost modelling

\emph{In the problem you're describing, one natural idea that springs to mind is
  to reduce it back to a graph problem by the edges of the original interaction
  graph as vertices of an "augmented interaction graph" whose vertex set
  consists of both the vertices and edges of the original graph, i.e. all
  entities that might have data associated to them. And two vertices of the
  augmented interaction graph are joined by an edge if either (a) both of them
  are vertices of the original interaction graph, and they are joined by an
  edge, or (b) one is an edge and the other is a vertex which is an endpoint of
  that edge. Pictorially, the augmented interaction graph is obtained by
  inserting a new vertex alongside each edge of the original interaction graph
  and turning the edge into a triangle with the new vertex as its third
  corner. It seems that this would work well assuming that queries which
  traverse an edge are likely to traverse its endpoints. -- rdk}

@@

>>>>

\emph{In the problem you're describing, one natural idea that springs to mind is
  to reduce it back to a graph problem by the edges of the original interaction
  graph as vertices of an ``augmented interaction graph" whose vertex set
  consists of both the vertices and edges of the original graph, i.e. all
  entities that might have data associated to them. And two vertices of the
  augmented interaction graph are joined by an edge if either (a) both of them
  are vertices of the original interaction graph, and they are joined by an
  edge, or (b) one is an edge and the other is a vertex which is an endpoint of
  that edge. Pictorially, the augmented interaction graph is obtained by
  inserting a new vertex alongside each edge of the original interaction graph
  and turning the edge into a triangle with the new vertex as its third
  corner. It seems that this would work well assuming that queries which
  traverse an edge are likely to traverse its endpoints. -- rdk}

@@

Let $Q$ be the query workload, where each query $q\in Q$ accesses a set of
attributes $q.A$ and traverses parts of the graph for the time range
$q.T=[q.t_s,q.t_e]$. We denote the set of all attributes as $A$. A given block
$B$, we denote its time range as $B.T$, which is the union of the time ranges
of its temporal neighborlists. Let $s(a)$ denote the size of an attribute $a$.
We also use the same notation for block size, that is $s(B)$ for a block $B$. 
We use $c(B)$ to denote the number of entries in the temporal neighborlists
within block $B$, that is the number of edges in the block.

Our goal is to create an overlapping partitioning of a block, denoted by
$\mathcal{P}(B)$. We call the partitions \emph{sub-blocks}. We have
$\bigcup_{B'\in \mathcal{P}(B)} B'.A = A$. We aim to find the function
$\mathcal{P}$ that minimizes the query I/O over $B$, while keeping the
relative storage overhead beyond a limit, say $1+\alpha$ times the original.
If we represent the query I/O as $L(\mathcal{P}(B))$ and the relative storage
overhead as $H(\mathcal{P}(B))$, our goal is to find:
\begin{equation}
\mathcal{P} \leftarrow \mbox{argmax}_{\{\mathcal{P}: H(\mathcal{P}(B)) < \alpha\}} L(\mathcal{P}(B))
\end{equation}

The storage overhead can be formalized as:
\begin{equation}
H(\mathcal{P}(B)) = (1-\frac{\sum_{a\in A} s(a)}{s(B)})\cdot |\mathcal{P}(B)| 
\end{equation}

The Query I/O can be formalized as:\\
TODO for Robert.\\
We need to use the workload characteristics for the blocks time range as
well as $\mathcal{P}(B)$ to formalize the query I/O.\\
